### 2024-07-10 Article/Article-english.tex
finish fourth chapter

    Take an~arbitrary character $f$. Since it is a~homomorphism, $f(1_G) = 1$
    where $1_G$ is a~unit in~the~group $G$. But then if $\alpha f = 0$, then
    $\alpha = \alpha \cdot 1 = \alpha \cdot f(1_G) = 0$ what proves the~base case.

@@

We will use the~existence of~\textit{Smith's normal form} for~integer matrices \cite{Smth60}.
Let $\mathrm{diag}^{m \times n}_r(x_1,\allowbreak \ldots,\allowbreak x_r)$ be a~diagonal matrix $\mathrm{diag}(x_1,\allowbreak \ldots,\allowbreak x_r)$
augmented (or cut off) from the~bottom right by~zeroes to a~matrix of size $m \times n$.

\medskip\noindent\textbf{Theorem} (on~the~existence of~Smith's normal form).\emph{
    $$
        \forall \alpha \in \mathbb{Z}^{m \times n}{:}\
        \exists \beta_1 \in \mathrm{GL}^m(\mathbb{Z}), \beta_2 \in \mathrm{GL}^n(\mathbb{Z}){:} \
        \exists \varepsilon_1, \ldots, \varepsilon_r \in \mathbb{Z} \setminus \{0\}{:} \
        \beta_1 \alpha \beta_2 = \mathrm{diag}^{m \times n}_r(\varepsilon_1, \ldots, \varepsilon_r),
    $$
    where $\varepsilon_1 \mid \varepsilon_2 \mid \ldots \mid \varepsilon_r$ (the so~called invariant factors) and~$r = \mathrm{rank}(\alpha)$.
}\medskip

@@

As in~the~case of~the~theory of~curves and surfaces, we say that a~subgroup of~an~algebraic torus
is \textit{parameterizable} if it can be represented as an~image of~some mapping. Similarly, a~subgroup
is \textit{monomially parameterizable} if it can be represented as an~image of~$\phi_\alpha$ for some matrix $\alpha$.

Since $\phi_\alpha$ is the~homomorphism, its image $\mathrm{Im}(\phi_\alpha)$ is a~subgroup in~$(K^\times)^n$.
Next we'll study two questions: whether $\mathrm{Im}(\phi_\alpha)$ is an~algebraic subgroup and whether
any algebraic subgroup is represented by~some $\phi_\alpha$.

@@

\medskip\noindent\textbf{Lemma~6.}\emph{
    $\phi_\alpha$ for $\alpha \in \mathbb{Z}^{n \times k}$ can be represented as some image $\mathrm{Im}(\phi_\beta)$
    if the~mappings $g \mapsto g^{\varepsilon_i}$ are surjective in~$G$ for all $\varepsilon_i$ из~$\mathrm{SNF}(\alpha)$.
}\medskip

@@

>>>>

    Take an~arbitrary character $f$. Since it is a~homomorphism, $f(1_G) = 1$
    where $1_G$ is an~identity element in~the~group $G$. But then if $\alpha f = 0$, then
    $\alpha = \alpha \cdot 1 = \alpha \cdot f(1_G) = 0$ what proves the~base case.

@@

We will use the~existence of~\textit{Smith normal form} for~integer matrices \cite{Smth60}.
Let $\mathrm{diag}^{m \times n}_r(x_1,\allowbreak \ldots,\allowbreak x_r)$ be a~diagonal matrix $\mathrm{diag}(x_1,\allowbreak \ldots,\allowbreak x_r)$
augmented (or cut off) from the~bottom right by~zeroes to a~matrix of size $m \times n$.

\medskip\noindent\textbf{Theorem} (on~the~existence of~Smith normal form).\emph{
    $$
        \forall \alpha \in \mathbb{Z}^{m \times n}{:}\
        \exists \beta_1 \in \mathrm{GL}^m(\mathbb{Z}), \beta_2 \in \mathrm{GL}^n(\mathbb{Z}){:} \
        \exists \varepsilon_1, \ldots, \varepsilon_r \in \mathbb{Z} \setminus \{0\}{:} \
        \beta_1 \alpha \beta_2 = \mathrm{diag}^{m \times n}_r(\varepsilon_1, \ldots, \varepsilon_r),
    $$
    where $\varepsilon_1 \mid \varepsilon_2 \mid \ldots \mid \varepsilon_r$ (the so~called invariant factors) and~$r = \mathrm{rank}(\alpha)$.
}\medskip

@@

As in~the~case of~the~theory of~curves and surfaces, we say that a~subgroup of~an~algebraic torus
is \textit{parameterizable} if it can be expressed as an~image of~some mapping. Similarly, a~subgroup
is \textit{monomially parameterizable} if it can be expressed as an~image of~$\phi_\alpha$ for some matrix $\alpha$.

Since $\phi_\alpha$ is the~homomorphism, its image $\mathrm{Im}(\phi_\alpha)$ is a~subgroup in~$(K^\times)^n$.
Next we'll study two questions: whether $\mathrm{Im}(\phi_\alpha)$ is an~algebraic subgroup and whether
any algebraic subgroup is expressed by~some $\phi_\alpha$.

@@

\medskip\noindent\textbf{Lemma~6.}\emph{
    $\phi_\alpha$ for $\alpha \in \mathbb{Z}^{n \times k}$ can be expressed as some image $\mathrm{Im}(\phi_\beta)$
    if the~mappings $g \mapsto g^{\varepsilon_i}$ are surjective in~$G$ for all $\varepsilon_i$ из~$\mathrm{SNF}(\alpha)$.
}\medskip

@@

\noindent{\it Proof.}
    Let $H = \ker(\phi_\beta)$. Without loss of generality, we assume that the~rows of~the~matrix $\beta$ are linearly independent
    over $Z$ (it is clear that the~rows expressed as a~linear combination of~the~other rows can be removed from the~matrix $\beta$
    without changing the~kernel).

    Let us write the~Smith normal form for $\beta$: $\beta = \beta_1^{-1} \varepsilon \beta_2^{-1}$.
    Since, according to~the~remark above, $\beta$ has full rank, the~matrix $\varepsilon$ has no zero rows.
    Furthermore, $\Pi(\beta) = 1$, so $|\omega_K(\varepsilon_i)| = 1$, that is, the~equations $z^{\varepsilon_i} = 1$
    have only $z = 1$ as a~solution.

    The~kernel of~$\varepsilon$ is given by~vectors of~the~form $(0,\allowbreak \ldots,\allowbreak 0,\allowbreak t_{k + 1},\allowbreak \ldots, t_{n})$.
    Consider a~matrix $\delta \in \mathbb{Z}^{n \times (n - k)}$ corresponding to~the~linear operator
    $$
        (t_1, \ldots, t_{n - k}) \mapsto (0, \ldots, 0, t_1, \ldots, t_{n - k}).
    $$

    Let also $\alpha = \beta_2 \delta$. We'll prove that $\mathrm{Im}(\phi_\alpha) = \ker(\phi_\beta)$.

    Indeed, by~definition of~$\delta$ it holds that $\delta$, $\varepsilon \delta = 0$, so $\phi_{\beta} \circ \phi_{\alpha} = \phi_{\beta_1^{-1} \varepsilon \beta_2^{-1} \beta_2 \delta} = \phi_{\beta_1^{-1} \varepsilon \delta} = \phi_{0} = 1$,
    that is, $\mathrm{Im}(\phi_{\alpha}) \subseteq \ker(\phi_{\beta})$.

    Conversely, let $z \in \ker(\phi_\beta)$. Then $\phi_{\beta_1^{-1}} (\phi_{\varepsilon \beta_2^{-1}}(z)) = \phi_\beta(z) = 1$,
    so $\phi_{\varepsilon}(\phi_{\beta_2^{-1}}(z)) = \phi_{\varepsilon \beta_2^{-1}}(z) = \phi_{\beta_1}(1) = 1$.
    From the~form of~the~matrix $\varepsilon$ we obtain that
    $
        \phi_{\beta_2^{-1}}(z) = (1,\allowbreak \ldots,\allowbreak 1,\allowbreak t_{k + 1},\allowbreak \ldots, t_n) = \phi_\delta(t_{k + 1}, \ldots, t_n)
    $; that is, $z = \phi_{\beta_2}(\phi_\delta(t_{k + 1}, \ldots, t_n)) = \phi_\alpha(t_{k + 1}, \ldots, t_n)$ for~some $t_j$.
    Thus, $z \in \mathrm{Im}(\phi_\alpha)$ and $\ker(\phi_\beta) \subseteq \mathrm{Im}(\phi_\alpha)$.
\hfill$\Box$\medskip

\medskip\noindent\textbf{Теорема 3.}\emph{
    The~number of~connected components of~$\ker(\phi_\beta) \subseteq ({\mathbb{C}^\times})^n$ is equal to~$\Pi(\beta)$.
}\medskip

\noindent{\it Proof.}
    Consider again the~Smith normal form of~$\beta$: $\beta$ = $\beta_1^{-1} \varepsilon \beta_2^{-1}$.
    $\phi_{\beta_1^{-1}}$ is isomorphism, so $\ker(\phi_{\beta}) = \ker(\phi_{\varepsilon \beta_2^{-1}})$.

    We denote the~rows of~$\beta_2^{-1}$ by~$b_i$. For a~vector $u \in \omega(\varepsilon)$ consider
    the~set $H_u = \{z \in ({\mathbb{C}^\times})^n \ | \ \forall 1 \leq i \leq r{:}\ z^{b_i} = u_i\}$.
    The~condition $\phi_\varepsilon(\phi_{\beta_2^{-1}}(z)) = 1$ is obviously equivalent to~the~condition
    $\exists u \in \omega(\varepsilon){:}\ z \in H_u$. Clearly, the~sets $H_u$ are disjunctive,
    so $\ker(\phi_{\beta})$ decomposes into a~disjunctive union:
    $
        \ker(\phi_{\beta}) = \bigsqcup_{u \in \omega(\varepsilon)} H_u.
    $
    There are exactly $\Pi(\beta)$ vectors $u \in \omega(\varepsilon)$, so it suffices to~show that
    each component $H_u$ is connected.

    Since, by~definition, $H_1 = \ker(\phi_{\delta \beta_2^{-1}})$ where $\delta = \mathrm{diag}^{k \times n}_r(1, \ldots, 1)$
    and $\Pi(\delta \beta_2^{-1}) = 1$, the~component $H_1$ is connected by~the~remark.

    Fix $u$ and consider the~matrix $\tau = \mathrm{diag}(1 / u_1, \ldots, 1 / u_r, 1, \ldots, 1)$
    and~the~mapping $\psi = \phi_{\beta_2} \circ \phi_\tau \circ \phi_{\beta_2^{-1}}$.
    $\psi$ is a~continuous bijection, moreover, $\psi^{-1} = \phi_{\beta_2} \circ \phi_{\tau^{-1}} \circ \phi_{\beta_2^{-1}}$.

    By~definition, $\phi_{\beta_2^{-1}}(\psi(z)) = \phi_{\tau \beta_2^{-1}}(z)$. So if $z \in H_u$,
    then $\psi(z)^{b_i} = z^{b_i} / u_i = u_i / u_i = 1$. Conversely, if $z \in H_1$, then $\psi^{-1}(z)^{b_i} = u_i z^{b_i} = u_i$.
    Thus, $\psi(H_u) = H_1$, that is, $H_u$ is homeomorphic to~$H_1$, but $H_1$ is connected.
\hfill$\Box$\medskip

As we saw before, any algebraic subgroup $H$ of the~torus $(K^\times)^n$ can be expressed as $\ker(\phi_\beta)$
for some $\beta$. Notice that $\Pi(\beta)$ does not depend on~the~choice of~$\beta$ for~the~group $H$.

\medskip\noindent\textbf{Theorem~4.}\emph{
    If $\mathrm{char}(K) = 0$ and $\ker(\phi_{\beta}) = \ker(\phi_{\beta'})$, then $\Pi(\beta) = \Pi(\beta')$.
}\medskip

\noindent{\it Proof.}
    Let us write down the~Smith normal forms: $\beta = \beta_1^{-1} \varepsilon \beta_2^{-1}$
    and~$\beta' = {\beta'}_1^{-1} \varepsilon' {\beta'}_2^{-1}$. So we have: $\ker(\phi_{\varepsilon \beta_2^{-1}}) = \ker(\phi_{\beta}) = \ker(\phi_{\beta'}) = \ker(\phi_{\varepsilon' {\beta'}_2^{-1}})$.

    Let $\varepsilon = \mathrm{diag}^{k \times n}_r(\varepsilon_1, \ldots, \varepsilon_r)$
    and~$\varepsilon' = \mathrm{diag}^{k' \times n}_{r'}(\varepsilon'_1, \ldots, \varepsilon'_r)$.
    We write $\delta = \mathrm{diag}^{k \times n}_r(1, \ldots, 1)$ and~$\delta' = \mathrm{diag}^{k' \times n}_{r'}(1, \ldots, 1)$.

    Since $\Pi(\delta \beta_2^{-1}) = \Pi(\delta' {\beta'}_2^{-1}) = 1$, both groups are parameterized
    according to~lemma~7 by~some $\phi_\alpha$ и~$\phi_{\alpha'}$ respectively.

    $\mathrm{Im}(\phi_\alpha) = \ker(\phi_{\delta \beta_2^{-1}}) \subseteq \ker(\phi_{\varepsilon \beta_2^{-1}}) = \ker(\phi_{\varepsilon' {\beta'}_2^{-1}})$,
    so $\phi_{\varepsilon' {\beta'}_2^{-1}} \circ \phi_\alpha = 1$. Using lemma~2, $\varepsilon' {\beta'}_2^{-1} \alpha = 0$.
    Multiplying both sides of~the~equality by~the~matrix $\mathrm{diag}(1 / \varepsilon'_1, \ldots, 1 / \varepsilon'_r, 1, \ldots, 1)$
    one obtains that $\delta' {\beta'}_2^{-1} \alpha = 0$; but this means that $\ker(\phi_{\delta \beta_2^{-1}}) = \mathrm{Im}(\phi_\alpha) \subseteq \ker(\phi_{\delta' {\beta'}_2^{-1}})$.
    Similarly we get the~converse inclusion. Thus, $\ker(\phi_{\delta \beta_2^{-1}}) = \ker(\phi_{\delta' {\beta'}_2^{-1}})$.

    Consider the~quotient group $\ker(\phi_\beta) / \ker(\phi_{\delta \beta_2^{-1}}) = \ker(\phi_{\beta'}) / \ker(\phi_{\delta' {\beta'}_2^{-1}})$,
    also called \textit{the~component group.} The~mapping $\phi_{\delta \beta_2^{-1}} : \ker(\phi_\beta) \rightarrow \omega_K(\varepsilon) \times \{ 1 \} \times \ldots \times \{ 1 \}$
    induces an~injective homomorphism from the~quotient. In addtion, in~the~theorem~3 the~bijection between the~components $H_u$
    was constructed (it generalises unchanged to~the~case of~an~arbitrary field $K$), and hence they are all nonempty.
    This means that $\phi_{\delta \beta_2^{-1}}$ is surjective, so the~induced mapping is also surjective.
    The~reasoning is similar for~the~group $\omega_K(\varepsilon')$. Thus, we have a~chain of~isomorphisms:
    $$
        \omega(\varepsilon) \cong \ker(\phi_\beta) / \ker(\phi_{\delta \beta_2^{-1}}) = \ker(\phi_{\beta'}) / \ker(\phi_{\delta' {\beta'}_2^{-1}}) \cong \omega(\varepsilon').
    $$

    Isomorphic groups have the~same order, and hence $\Pi(\beta) = |\omega_K(\varepsilon)| = |\omega_K(\varepsilon')| = \Pi(\beta')$.
\hfill$\Box$\medskip

Finally, for any algebraic subgroup $H$ of~the~torus we can define a~number $\Pi(H)$ equal to~$\Pi(\beta)$
for any $\beta$ such that $H = \ker(\phi_\beta)$.

Now for any algebraic subgroup $H \subseteq (\mathbb{C}^\times)^n$ we defined \textit{the~identity component} $H^\circ$
as a~connected component containing an~identity element of~the~group $H$. It follows from theorem~3
that $\Pi(H^\circ) = 1$. Thus, the~following statements are proved.

\medskip\noindent\textbf{Theorem~5.}\emph{
    An~algebraic subgroup $H$ of~the~torus $(K^\times)^n$ is parameterizable if and only if $\Pi(H) = 1$.
}

\medskip\noindent\textbf{Consequence.}\emph{
    An~algebraic subgroup $H$ of~the~torus $(\mathbb{C}^\times)^n$ is parameterizable if and only if it is connected.
}

\medskip\noindent\textbf{Consequence.}\emph{
    Every algebraic subgroup $H$ of~the~torus $(\mathbb{C}^\times)^n$ contains a~parameterizable subgroup $H^\circ$ of~the~same dimension.
}\medskip

\noindent{\it Proof.}
    All components of~$H$ are homeomorphic, what immediately proves the~theorem.
\hfill$\Box$\medskip
